---
title: "Tidy data"
author: "Sara Pedro"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)

# to call stats filter - stats::filter()
# the code above does not let warnings and messages show on the markdown version
```


# Reformat catch data

* Remove "All" and "notes" columns
* Create "Species" column
    - move from wide to long
* General QA

    ## Data citation

[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.] (https://knb.ecoinformatics.org/view/df35b.304.2)



```{r}

# Read data from a url - from the an online data bank

catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)

head(catch_original)

```

Remove the "All" and "notesRegCode" columns

ctrl+shit+m : pipe operator shortcut %>% 

```{r}

catch_long <- catch_original %>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% # or select(-All,-notesRegCode)
  gather(key = species, value = catch, -Region, -Year) # Move from wide to long; Region and                                                             Year are not going to be gathered                                                               into the species and catch columns


head(catch_long)

```


Opposite code of gather is spread - gets data set back from long to wide

```{r}
catch_wide <- catch_long %>% 
  spread (key = species, value = catch)

head(catch_wide)

```

## Clean up our data

* rename catch to catch_thousands
* change the catch column to numeric
* create a new catch column in units '#' of fish

```{r}
catch_clean <- catch_long %>% 
  rename(catch_thousands = catch) %>% # rename column catch
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% # replace                                      the "I" mistake in row 401 (error found below) for 1. Keep                                       the other rows
  mutate(catch_thousands = as.numeric(catch_thousands)) %>%
  mutate(catch = catch_thousands*1000) %>% # add new column catch by multiplying                                                            catch_thousands by 1000
  select(-catch_thousands) #if we don't want this column

head(catch_clean)

```



```{r, eval = FALSE, echo = FALSE}
#Investigate why I am getting NAs - look for where as.numeric failed 

test <- as.numeric(catch_long$catch)
i <- which(is.na(test)==TRUE) # which values in the test are NAs when tured to numeric
i
catch_long[i,]

```

## Split - Apply - Combine

* calculate mean catch by species

```{r}

species_mean <- catch_clean %>% 
  group_by(species, Year) %>%  # needs to be used with the functions summarize or mutate
  summarise(catch_mean = mean(catch), 
            num_obs = n()) # calculate mean and number of observations

year_mean <- catch_clean %>% 
  group_by(Year) %>% 
  summarise(catch_mean = mean(catch), stand_dev = sd(catch)) %>% 
  arrange(-Year) %>%  # arrange by highest to lowest (only for numeric). Could use desc() to                            change the order of species (categorical or numeric)
  filter(Year <= 1990 & Year >= 1960) # only year above 1990

head(species_mean)
head(year_mean)

```


# Join the Region Definitions

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"), stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea) #selects only the columns we are interested in

head(region_defs)
```

Left join

```{r}
catch_joined <- left_join(catch_clean, region_defs, by = c("Region" = "code"))

# could also do: catch_joined <- catch_clean %>%
                # left_join(region_defs)

head(catch_joined)
```


# Miscelaneous Functions

Separate and unite

Using some dummy data 
 
```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)
```


Separate - one column to different columns

```{r}
dates_split <- dates_df %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", remove = F) #remove equals false                                                                 will keep the original column

head(dates_split)

  
```


Unite


```{r}
dates_unite <- dates_split %>% 
  unite(date_iso, year, month, day, sep = "-")

head(dates_unite)
```




