---
title: "Geospatial Analysis"
author: "Sara Pedro"
date: "February 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(sf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(leaflet)
library(scales)
```

Data from:
[Jeanette Clark, Sharis Ochs, Derek Strong, and National Historic Geographic Information System. 2018. Languages used in Alaskan households, 1990-2015. Knowledge Network for Biocomplexity. doi:10.5063/F11G0JHX]

```{r}
ak_shp_sf <- read_sf("shapefiles/ak_regions_simp.shp")

head(ak_shp_sf)
# if the geometry column is in sf format, R knows how to read it with this package

```

```{r}
plot(ak_shp_sf)

```

```{r}
st_crs(ak_shp_sf) #indicates the coordinate reference system in this data
```
 
Can check spatial references for projections online 
 
```{r}
ak_shp_sf <- ak_shp_sf %>% 
  st_transform(crs = 3338) #coordinate reference system 

```


## Spatial joins
Join the population data with the regional information

```{r}
pop <- read.csv("shapefiles/alaska_population.csv")
```

Turn our population data.frame with latitude and longitude coordinates into an sf object. Let’s assume that the coordinate system used to reference the latitude longitude coordinates is WGS84, which has a crs number of 4236.

```{r}
# Convert foreign object to a sf object

pop_sf <- st_as_sf(pop,
                   coords = c("lng", "lat"),
                   crs = 4326,
                   remove = F) %>% 
# we created the geometry column readable in sf
  st_transform(crs = 3338)
# and now we converted it to the projection we want

```

Spatial joint to assign the regions to the population dataframe

```{r}
pop_joined_sf <- st_join(pop_sf, ak_shp_sf, join = st_within)
# st_join is a spatial inner of left join. In addition, we want to find what region each city falls within, so we will use st_within

plot(pop_joined_sf["region"])

```

## Group by and summarise

Getting population in the ak_shp_sf data set

Group by region

```{r}
pop_region <- pop_joined_sf %>%
  as.data.frame() %>% 
#group_by for sf objects (see ?group_by_sf) works a bit differently than the regular goup_by function
  group_by(region) %>% 
  summarise(total_pop = sum(population))

head(pop_region)
  
```

```{r}
ak_pop_sf <- left_join(ak_shp_sf, pop_region)

plot(ak_pop_sf["total_pop"])
```

Group by management area

```{r}
ak_mgmt <- ak_pop_sf %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop = sum(total_pop), do_union = F)

plot(ak_mgmt["total_pop"])

```

Write shape files

```{r}
write_sf(ak_shp_sf, "shapefiles/ak_regions_population.shp") 
```

Plotting maps with ggplot - allows to add a bunch of layers

```{r}
ggplot(ak_pop_sf) +
  geom_sf(aes(fill = total_pop)) +
  theme_bw() +
  labs(fill = "Total population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma) 


```

Plot more layers

```{r}
ggplot() +
  geom_sf(data = ak_pop_sf, aes(fill = total_pop)) +
  geom_sf(dat = pop_sf, size = 0.5) +
  # plots the location of the points
  theme_bw() +
  labs(fill = "Total population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma) 

```


## Interactive maps with leaflet


```{r}
# Creating a projection object in leaflet

epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))


```


```{r}
# transform to geographic coordinate sytem - leaflet needs this format
ak_crs <- st_transform(ak_pop_sf, crs = 4326)
pop_crs <- st_transform(pop_sf, crs = 4326)

```


```{r}
#Creates a pallete of colors to match our data
pal <- colorNumeric(palette = "Reds", domain = ak_crs$total_pop)

#leaflet plot
m <- leaflet(options = leafletOptions(crs = epsg3338)) %>% 
  addPolygons(data = ak_crs,
              fillColor = ~pal(total_pop),
              weight = 1,
              color ="black",
              label = ~region,
              fillOpacity = 1) %>% 
  addCircleMarkers(data = pop_crs,
                   lat = ~lat,
                   lng = ~lng,
                   radius = 2,
                   fillColor = "black",
                   fillOpacity = 1,
                   weight = 0.25,
                   color = "black",
                   label = ~city)

m
```


